---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(dplyr)
library(corrplot)
library(MASS)
library(caret)
library(caTools)
library(rpart)
library(rpart.plot)
library(gbm)
library(ggplot2)
library(randomForest)
library(party)
library(e1071)
library(vip)
```

```{r}
#Load datasets
ds1 <- read.csv("dataset1_2.csv")
ds2 <- read.csv("dataobserved_2.csv")

#Rename columns and remove unused row for dataset 1
colnames(ds1)[1] <- "ID"
colnames(ds1)[2] <- "Treatment"
colnames(ds1)[3] <- "Age"

#Rename columns for data set 2
colnames(ds2)[1] <- "ID"
colnames(ds2)[2] <- "Weight Change"
colnames(ds2)[3] <- "Visit Number"
```

```{r}
#Merge data set 1 and 2 on ID column. 
#Data includees 5 columns (ID(Unique Identifier), Weight Change, Visit Number, Treated (1 = Treated, 0 = Not Treated), Age)
wl <- merge(ds2, ds1, by.x = "ID", by.y = "ID")
str(wl)
```  

```{r}
#Change data type to best fit model
wl$Age <- as.integer(wl$Age)
wl$Treatment <- as.factor(wl$Treatment)

#Correlation check
#No correlation between Weight Change and the number of visits. 
corwl <- wl[,-4]
wl_mat <- cor(corwl)
corrplot(wl_mat, method = "number")
```

```{r}
#Normally distribution
hist(wl$`Weight Change`)
```


```{r}
#  Hypothesis testing
#Use 2 different Stats test

#t-Test 
#Ho:Treatment has no impact on weight 
#Ha: Treatment does have an impact on weight
#p-value == 2.2e-16
#Alpha is < 0.05 so we reject Ho
t.test(wl$`Weight Change`~ wl$Treatment)
```


```{r}
#Anova
#Ho:Treatment has no impact on weight 
#Ha: Treatment does have an impact on weight
#p-value == <2e-16
##Alpha is < 0.05 so we reject Ho
aov_fit <- aov(wl$`Weight Change`~wl$Treatment + wl$`Visit Number`+ wl$Age,data = wl)
summary(aov_fit)
```

```{r}
#Subset data into train and test set for ML model
set.seed(123)
sample <- sample.split(wl, SplitRatio = .70)
train <- subset(wl, sample == TRUE)
test <- subset(wl, sample == FALSE)

#Regression Tree model
wl_dc <- rpart(`Weight Change`~ Age + `Visit Number` + Treatment ,data = train, method = "anova")
#Treatment is the most important variable
rpart.plot(wl_dc)
#Actual weight loss vs predicted predicted
predicted_wl <- predict(wl_dc, test, method = "anova")
dc_df <- cbind(test, predicted_wl)
#Predited weight values 
#From or results we can infer that out model predicted weight loss with good accuracy. 
#Our model proved that if a patient received treatment they were more likely to lose weight on the other hand 
#if a patient did not receive treatment they were less likely to lose weight and in many cases gained weight.
```

```{r}
#Feature importance 
#Determine which variable was most important
vip_fit <- vip(wl_dc, num_features = 3, bar = FALSE)
print(vip_fit)
```

```{r}
#Subset original data frame into 5 smaller data frames based on visit number.
visit_1 <- wl[wl$`Visit Number` == 1,]      
visit_2 <- wl[wl$`Visit Number` == 2,]      
visit_3 <- wl[wl$`Visit Number` == 3,]      
visit_4 <- wl[wl$`Visit Number` == 4,]  
visit_5 <- wl[wl$`Visit Number` == 5,]
```

```{r}
#Visit 1 weight 
ggplot(data = v1, aes(x = `Weight Change`, y = Age)) +
geom_point(aes(color = Treatment), shape = 1, alpha = 1, size = 2) + 
labs(title="Weight Loss Based on Treatment or No Treatment", y="Age", x="Weight Loss") +
scale_color_discrete(name="Treatment",
                    breaks=c("0", "1"),
                    labels=c("0 = Not Treated", "1 = Treated "))
```

```{r}
#Visit 2 weight 
ggplot(data = v2, aes(x = `Weight Change`, y = Age)) +
geom_point(aes(color = Treatment), shape = 1, alpha = 1, size = 2) +
labs(title="Weight Loss Based on Treatment or No Treatment", y="Age", x="Weight Loss") +
scale_color_discrete(name="Treatment",
                    breaks=c("0", "1"),
                    labels=c("0 = Not Treated", "1 = Treated "))
```

```{r}
#Visit 3 weight 
ggplot(data = v3, aes(x = `Weight Change`, y = Age)) +
geom_point(aes(color = Treatment), shape = 1, alpha = 1, size = 2) +
labs(title="Weight Loss Based on Treatment or No Treatment", y="Age", x="Weight Loss") +
scale_color_discrete(name="Treatment",
                    breaks=c("0", "1"),
                    labels=c("0 = Not Treated", "1 = Treated "))
```

```{r}
#Visit 4 weight 
ggplot(data = v4, aes(x = `Weight Change`, y = Age)) +
geom_point(aes(color = Treatment), shape = 1, alpha = 1, size = 2) +
labs(title="Weight Loss Based on Treatment or No Treatment", y="Age", x="Weight Loss") +
scale_color_discrete(name="Treatment",
                    breaks=c("0", "1"),
                    labels=c("0 = Not Treated", "1 = Treated"))
```

```{r}
#Visit 5 weight 
ggplot(data = v5, aes(x = `Weight Change`, y = Age)) +
geom_point(aes(color = Treatment), shape = 1, alpha = 1, size = 2) +
labs(title="Weight Loss Based on Treatment or No Treatment", y="Age", x="Weight Loss") +
scale_color_discrete(name="Treatment",
                    breaks=c("0", "1"),
                    labels=c("0 = Not Treated", "1 = Treated "))
```

#Regression tree analysis was chosen because it is a reliable method of identifying which variables have impact on a topic of #interest which in this case was the change in weight. The process of performing a regression allows you to confidently #determine which factors matter most, which factors can be ignored, and how these factors influence each other. 

#Prior to building our model we check for correlation with all the variables. Unfortunately, there was no correlation. After #model building we determined there is more so causation than correlation. 

#Statistical testing was performed using a t-test and a two-way Anova. In both cases we were able to reject the null hypothesis #and accept the alternative which was our original assumption that treatment led to weight loss. 

#There is definitely a treatment effect on weight over time. As displayed by the above visualizations. After building our model #we wanted further proof that treatment over time affected weight. We divided our data set into patient visits (1-5) to track #patient weight loss or patient weight gained over time. Our visuals provide further evidence that patients who received treatment saw a significant difference in weight loss over the 5 visits.
